
```{r include = F, eval = T}
source("~/bio/code/mpip/dex-stim-human-array/code/integrative/util.R")
```

***
# __Identification of independent meQTLs__
***

## Methodology

The number of independent SNPs per cis-region was identified by LD pruning and clumping the SNPs using the “clump” command in PLINK (using distance < __200 kb__ and __r2 > 0.2__ as setting). Each SNP forms a SNP bin, by aggregating all other SNPs into bins by the tag SNP at r2 > 0.2 and distance < 200 kb, such that all SNPs within a given bin were correlated to the current SNP, but to any other SNP. 

```{r include = F, eval = T}
cbPalette <- c( "#0072B2", "#009E73", "#E69F00")
```

```{r, include = F, eval = T}
library(data.table)
out.dir.pre  <- "~/bio/code/mpip/dex-stim-human-array/output/data/integrative/matrixEQTL/meqtls/region_wise_independent_snps/"

meqtl.dex.fn   <- paste0(out.dir.pre, "me-qtl_cis_indp_rw_dex_fdr_005.csv")
meqtl.veh.fn   <- paste0(out.dir.pre, "me-qtl_cis_indp_rw_veh_fdr_005.csv")
meqtl.delta.fn <- paste0(out.dir.pre, "me-qtl_cis_indp_rw_delta_fdr_005.csv")

col.names <-  c("CpG_ID", "SNP", "beta", "t-stat", "p-value", "fdr")

ind.meqtl.dex.df   <- fread(meqtl.dex.fn) #, col.names = col.names)
ind.meqtl.veh.df   <- fread(meqtl.veh.fn, col.names = col.names)
ind.meqtl.delta.df <- fread(meqtl.delta.fn) #, col.names = col.names)
```

```{r, include = F, eval = T}
meqtl.all.full.df <- rbind(ind.meqtl.dex.df[, treatment := "dex"],
                           ind.meqtl.veh.df[, treatment := "veh"],
                           ind.meqtl.delta.df[, treatment := "delta"])

meqtl.all.full.df <- meqtl.all.full.df[!is.na(meqtl.all.full.df$fdr),]
meqtl.all.full.df[["meQTL_ID"]] <- paste(meqtl.all.full.df$SNP, meqtl.all.full.df$CpG_ID, sep = "-")
meqtl.all.full.df <- setDT(meqtl.all.full.df)
```

```{r include = F, eval = T}
freq <- fread("/Users/anastasiia_hry/bio/code/mpip/dex-stim-human-array/data/snps/imputed_qc/from_janine/qc/freq.afreq")
```

## Significant Hits, FDR < 0.05

```{r, include = F, eval = T}
sign.hits.df <- data.table(c(ind.meqtl.delta.df$fdr  %>%  length(), 
                             ind.meqtl.delta.df$CpG_ID %>% unique() %>% length(),
                             ind.meqtl.delta.df$SNP %>% unique() %>% length()),
                           c(ind.meqtl.dex.df$fdr  %>% length(), 
                             ind.meqtl.dex.df$CpG_ID %>% unique() %>% length(),
                             ind.meqtl.dex.df$SNP %>% unique() %>% length()),
                           c(ind.meqtl.veh.df$fdr  %>% length(), 
                             ind.meqtl.veh.df$CpG_ID %>% unique() %>% length(),
                             ind.meqtl.veh.df$SNP %>% unique() %>% length()))

sign.hits.df <- as.data.frame(sign.hits.df)
rownames(sign.hits.df) <- c("cis-meQTLs", "CpG", "SNP")
colnames(sign.hits.df) <- c("delta", "dex", "veh")
```

```{r out.width = "95%", include = T, eval = T}
sign.hits.df %>% as.matrix() %>% reshape2::melt(var.names = c("ID", "Treatment"), value.name = "cnt") %>%
  ggplot(aes(y = cnt, x = Var2, fill = as.factor(as.numeric(Var2)))) +
    geom_col() +
    geom_text(aes(label = comma(cnt, accuracy = 1L), y = cnt), 
                  stat = "identity", 
                  vjust = -0.2, size = 2.5) + 
    scale_y_continuous(labels = scientific) +
    # scale_fill_viridis(discrete = TRUE, alpha = 0.9, option = "C") +
    labs(title = " ", y = "Count", x = " ") + 
    facet_wrap(~ Var1, ncol = 3) +
    theme(legend.position = "none", 
          legend.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(size = 10),
          axis.title = element_text(size = 8),
          axis.title.x = element_blank()) +
  scale_fill_manual(values = cbPalette)
```

## Upset plots{.tabset .tabset-fade .tabset-pills}

### meQTL level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.meqtls <- list(delta = meqtl.all.full.df[treatment == "delta", meQTL_ID], 
                      dex = meqtl.all.full.df[treatment == "dex", meQTL_ID], 
                      veh = meqtl.all.full.df[treatment == "veh", meQTL_ID])

upset(fromList(meqtls.meqtls), 
      sets = c("delta", "dex", "veh"),
      nsets = 3, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, meQTL", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = c( "#0072B2", "#009E73", "#E69F00"), 
      order.by = "freq")
```

### SNP level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.snps <- list(delta = ind.meqtl.delta.df$SNP, 
                    dex = ind.meqtl.dex.df$SNP, 
                    veh = ind.meqtl.veh.df$SNP)

upset(fromList(meqtls.snps), 
      sets = c("delta", "dex", "veh"),
      nsets = 3, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, SNP", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = c("#0072B2", "#009E73", "#E69F00"), 
      order.by = "freq")
```

### CpG level

```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.cpg <- list(delta = ind.meqtl.delta.df$CpG_ID, 
                   dex = ind.meqtl.dex.df$CpG_ID, 
                   veh = ind.meqtl.veh.df$CpG_ID)

upset(fromList(meqtls.cpg), 
      sets = c("delta", "dex", "veh"),
      nsets = 3, 
      nintersects = 10, 
      mainbar.y.label = "Number of intersections, CpG", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = c("#0072B2", "#009E73", "#E69F00"), 
      order.by = "freq",
      )
```

## Scatter plot

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05
plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")
GetScatterPlot2(meqtl.all.full.df, plot.title = plot.title, cbPalette = cbPalette)
```

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05

plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")

cbPalette <- c( "#0072B2", "#009E73", "#E69F00", "#F0E442", "#D55E00", "#CC79A7", "#56B4E9", "#999999")

ggplot(meqtl.all.full.df, aes(x = `t-stat`, y = -log10(fdr), shape = treatment, color = treatment)) +
  geom_point(alpha = 1.5, size = 1.2) +
  scale_x_continuous(labels = scientific) +
  geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
  geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "t-stat, MatrixEQTL",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```

## Only veh and delta

```{r include = F, eval = T}
cbPalette <- c( "#E69F00",  "#0072B2")
```

```{r out.width = "95%", include = T, eval = T}
sign.hits.df[, c("veh", "delta")] %>% as.matrix() %>% 
  reshape2::melt(var.names = c("ID", "Treatment"), value.name = "cnt") %>%
  ggplot(aes(y = cnt, x = Var2, fill = as.factor(as.numeric(Var2)))) +
    geom_col() +
    geom_text(aes(label = comma(cnt, accuracy = 1L), y = cnt), 
                  stat = "identity", 
                  vjust = -0.2, size = 2.5) + 
    scale_y_continuous(labels = scientific) +
    # scale_fill_viridis(discrete = TRUE, alpha = 0.9, option = "C") +
    labs(title = " ", y = "Count", x = " ") + 
    facet_wrap(~ Var1, ncol = 3) +
    theme(legend.position = "none", 
          legend.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(size = 10),
          axis.title = element_text(size = 8),
          axis.title.x = element_blank()) +
  scale_fill_manual(values = cbPalette)
```

## Upset plots{.tabset .tabset-fade .tabset-pills}

### meQTL level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.meqtls <- list(delta = meqtl.all.full.df[treatment == "delta", meQTL_ID], 
                      veh = meqtl.all.full.df[treatment == "veh", meQTL_ID])

upset(fromList(meqtls.meqtls), 
      sets = c("delta", "veh"),
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, meQTL", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = c("#0072B2", "#E69F00"), 
      order.by = "freq")
```

### SNP level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.snps <- list(delta = ind.meqtl.delta.df$SNP, 
                    veh = ind.meqtl.veh.df$SNP)

upset(fromList(meqtls.snps), 
      sets = c("delta", "veh"),
      nsets = 3, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, SNP", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color =  c("#0072B2", "#E69F00"),
      order.by = "freq")
```

### CpG level

```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.cpg <- list(delta = ind.meqtl.delta.df$CpG_ID, 
                   veh = ind.meqtl.veh.df$CpG_ID)

upset(fromList(meqtls.cpg), 
      sets = c("delta", "veh"),
      nsets = 3, 
      nintersects = 10, 
      mainbar.y.label = "Number of intersections, CpG", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color =  c("#0072B2", "#E69F00"),
      order.by = "freq",
      )
```

## Scatter plot

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05
plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")
cbPalette  <- c( "#0072B2", "#E69F00")
GetScatterPlot2(meqtl.all.full.df[treatment %in% c("veh", "delta")], plot.title = plot.title, cbPalette = cbPalette)
```

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05

plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")

ggplot(meqtl.all.full.df[treatment %in% c("veh", "delta")], 
       aes(x = `t-stat`, y = -log10(fdr), shape = treatment, color = treatment)) +
  geom_point(alpha = 1.5, size = 1.2) +
  scale_x_continuous(labels = scientific) +
  geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
  geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "t-stat, MatrixEQTL",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```
