---
title: "cis-meQTLs Analysis"
subtitle: "Dexamethasone-Stimulated Human Array Project"
author: 
  - name: Anastasiia Hryhorzhevska
    email: anastasiia_hry@psych.mpg.de
    url: https://github.com/ahryho
    affiliation: Max Planck Institute of Psychiatry
    affiliation_url: https://www.psych.mpg.de/2664393/medizinische-genomforschung
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    css: style.css
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
options(digits = 4, width = 100, stringsAsFactors = T)
knitr::opts_chunk$set(echo = TRUE,
                      tidy.opts = list(width.cutoff = 100),
                      tidy=TRUE,
                      fig.pos = "H",
                      dpi = 200,
                      warning = FALSE, 
                      message = FALSE, 
                      cache = TRUE, 
                      cache.lazy = FALSE)

# define DT as the standard printing method for data.frame
library(knitr)
library(DT)
knit_print.data.frame <- function(x, ...) {
  knit_print(DT::datatable(x,
                           filter="top",
                           rownames = FALSE,
                           extensions = "FixedColumns",
                           options = list(
                             scrollX = TRUE,
                             pageLength = 5
                             
                           )), ...)
}

registerS3method("knit_print", "data.frame", knit_print.data.frame)

library(data.table)
library(dplyr)
```

```{r}
setwd("~/bio/code/mpip/dex-stim-human-array/")
source("code/integrative/util.R")
```

```{r load-and-prepare-pub-data, eval=F}
# Load publicly available data from Josine L.Min at el. 2021

# mqtl_data <- fromJSON("http://api.godmc.org.uk/v0.1/assoc_meta/cpg/cg17242362", simplifyVector = FALSE)
 
public.meqtl.df <- fread("data/public_data/GoDMC_mQTL/assoc_meta_all.csv")
public.meqtl.cis.sub.df <- public.meqtl.df[cistrans == T, .(cpg, snp, pval)] # 71'616'458
public.meqtl.cis.sub.df[["meQTL_ID"]] <- paste(public.meqtl.cis.sub.df$cpg, public.meqtl.cis.sub.df$snp, sep = "-")

fwrite(public.meqtl.cis.sub.df, 
       "data/public_data/GoDMC_mQTL/assoc_meta_cis_meqtl.csv",
       quote = F, row.names = F, sep = ";")

fwrite(x = as.data.frame(public.meqtl.cis.sub.df$meQTL_ID), 
       "data/public_data/GoDMC_mQTL/assoc_meta_cis_meqtl_lst.csv",
       quote = F, row.names = F, sep = ";")
```

```{r load-public-meqtls, eval = T}
public.meqtl.cis.sub.df <- fread("data/public_data/GoDMC_mQTL/assoc_meta_cis_meqtl.csv")
  # fread("data/public_data/GoDMC_mQTL/assoc_meta_cis_meqtl_lst.csv")
```

```{r load-veh-meqtls, eval = T}
col.names    <- c("SNP", "CpG_ID", "beta", "t-stat", "p-value", "FDR")
meqtl.veh.df <- fread("output/data/integrative/matrixEQTL/meqtls/me-eqtl_cis_results_06122021/me-qtl_cis_result_veh_fdr_005.csv", col.names = col.names) 

meqtl.veh.nom.df <- fread("output/data/integrative/matrixEQTL/meqtls/me-eqtl_cis_results_06122021/me-qtl_cis_result_veh.csv", col.names = col.names) 
```

```{r load-dnam-veh}
methyl.beta.mtrx <- LoadMethylBeta("veh")
all.veh.cpgs     <- methyl.beta.mtrx$CpG_ID
```

```{r load-snp-data, eval = F}
# Genotype data

snp.loc.df <- fread("data/integrative/matrixEQTL/snp_locations.csv")

head(snp.loc.df)
dim(snp.loc.df)

snp.loc.df[["SNP_ID"]] <- paste0("chr", snp.loc.df$chr, ":", snp.loc.df$pos, ":", "SNP")
```

# Join baseline meQTL db with SNP ID
```{r}
meqtl.veh.df  <- left_join(meqtl.veh.df, snp.loc.df)
meqtl.veh.lst <- paste(meqtl.veh.df$CpG_ID, meqtl.veh.df$SNP_ID, sep = "-") # 6'944'026
meqtl.veh.lst
```

# Overlap at meQTL level: significant meQTLs @ FDR <= 0.05

```{r}
(intersect.meqtls <- intersect(meqtl.veh.lst, public.meqtl.cis.sub.df$meQTL_ID)) # 2'851'678

length(intersect.meqtls) / length(meqtl.veh.lst) # 41.1 %
```

# Overlap at meCpG level: significant meQTLs @ FDR <= 0.05
 
```{r}
veh.mecpgs <- meqtl.veh.df$CpG_ID %>% unique() # 137'537
pub.mecpgs <- public.meqtl.cis.sub.df$cpg %>% unique() # 232'477

intersect.all.veh.pub.cpgs <- intersect(all.veh.cpgs, pub.mecpgs) # 203'696
intersect.mecpgs.fdr       <- intersect(veh.mecpgs, intersect.all.veh.pub.cpgs) #53'891

length(intersect.mecpgs) / length(veh.mecpgs) # 39.18%
length(intersect.mecpgs) / length(intersect.all.veh.pub.cpgs) # 26.46 %
```

# Overlap at meCpG level: nominal significant meQTLs @ p <= 0.05
```{r}
veh.nom.mecpgs        <- meqtl.veh.nom.df$CpG_ID %>% unique() # 736'762
intersect.mecpgs.nom  <- intersect(veh.nom.mecpgs, intersect.all.veh.pub.cpgs) # 202'972

length(intersect.mecpgs.nom) / length(intersect.all.veh.pub.cpgs) # 99.64 %
```

# Direction of meQTLs: intersection (veh meQTLs nom; public)
```{r}
public.meqtl.nom.intersect.df <- public.meqtl.cis.sub.df[cpg %in% intersect.mecpgs.nom] # 59'400'674
public.meqtl.nom.intersect.df <- unique(public.meqtl.nom.intersect.df) # 59'400'674
```
