***
# __Significant independent meQTLs by region-wise approach. Visualisation__
***

```{r, include = F, eval = T}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(scales)
library(factoextra)
library(viridis)
library(knitr)
```

```{r eval = F, include = F}

### Independent SNPs
library(dplyr)

out.dir.pre  <- "~/bio/code/mpip/dex-stim-human-array/output/data/integrative/matrixEQTL/"

meqtl.dex.fn   <- paste0(out.dir.pre, "me-qtl_cis_result_dex_fdr_005.csv")
meqtl.veh.fn   <- paste0(out.dir.pre, "me-qtl_cis_result_veh_fdr_005.csv")
meqtl.delta.fn <- paste0(out.dir.pre, "me-qtl_cis_result_delta_fdr_005.csv")

col.names <-  c("SNP", "CpG_ID", "beta", "t-stat", "p-value", "fdr")

meqtl.dex.df   <- fread(meqtl.dex.fn, col.names = col.names)
meqtl.veh.df   <- fread(meqtl.veh.fn, col.names = col.names)
meqtl.delta.df <- fread(meqtl.delta.fn, col.names = col.names)

meqtl.delta.cpg.snp.ass     <- paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_delta_ind_cpg_snp_associations.txt")
meqtl.delta.cpg.snp.ass.lst <- fread(meqtl.cpg.snp.ass) 

meqtl.cpg.snp.ass     <- paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_veh_ind_cpg_snp_associations_v03.txt")
meqtl.cpg.snp.ass.lst <- fread(meqtl.cpg.snp.ass) 
meqtl.cpg.snp.ass.lst <- rbind(meqtl.cpg.snp.ass.lst, meqtl.delta.cpg.snp.ass.lst) %>% unique()
meqtl.veh.ind.rw.df   <- left_join(meqtl.cpg.snp.ass.lst, meqtl.veh.df)

fwrite(meqtl.veh.ind.rw.df,
       paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_indp_rw_veh_with_delta_fdr_005.csv"),
       quote = F, row.names = F, sep = ";")

meqtl.cpg.snp.ass     <- paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_dex_ind_cpg_snp_associations.txt")
meqtl.cpg.snp.ass.lst <- fread(meqtl.cpg.snp.ass) 
colnames(meqtl.cpg.snp.ass.lst) <- c("CpG_ID", "SNP")
meqtl.cpg.snp.ass.lst <- rbind(meqtl.cpg.snp.ass.lst, meqtl.delta.cpg.snp.ass.lst) %>% unique()
meqtl.dex.ind.rw.df <- left_join(meqtl.cpg.snp.ass.lst, meqtl.dex.df)

fwrite(meqtl.dex.ind.rw.df,
       paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_indp_rw_dex_with_delta_fdr_005.csv"),
       quote = F, row.names = F, sep = ";")
```

```{r include = F, eval = T}
out.dir.pre  <- "~/bio/code/mpip/dex-stim-human-array/output/data/integrative/matrixEQTL/region_wise_independent_snps/"

# meqtl.dex.fn   <- paste0(out.dir.pre, "me-qtl_cis_indp_result_dex_with_delta_fdr_005.csv")
# meqtl.veh.fn   <- paste0(out.dir.pre, "me-qtl_cis_indp_result_veh_with_delta_fdr_005.csv")
# meqtl.delta.fn <- paste0(out.dir.pre, "me-qtl_cis_indp_result_delta_fdr_005.csv")

meqtl.dex.fn   <- paste0(out.dir.pre, "me-qtl_cis_indp_rw_dex_with_delta_fdr_005.csv")
meqtl.veh.fn   <- paste0(out.dir.pre, "me-qtl_cis_indp_rw_veh_with_delta_fdr_005.csv")
meqtl.delta.fn <- paste0(out.dir.pre, "me-qtl_cis_indp_rw_delta_fdr_005.csv")

col.names <-  c("CpG_ID", "SNP", "beta", "t-stat", "p-value", "fdr")

ind.meqtl.dex.df   <- fread(meqtl.dex.fn, col.names = col.names)
ind.meqtl.veh.df   <- fread(meqtl.veh.fn, col.names = col.names)
ind.meqtl.delta.df <- fread(meqtl.delta.fn, col.names = col.names)
```

```{r, include = F, eval = T}
meqtl.all.full.df <- rbind(ind.meqtl.dex.df[, treatment := "dex"],
                           ind.meqtl.veh.df[, treatment := "veh"],
                           ind.meqtl.delta.df[, treatment := "delta"])

meqtl.all.full.df <- meqtl.all.full.df[!is.na(meqtl.all.full.df$fdr),]
meqtl.all.full.df[["meQTL_ID"]] <- paste(meqtl.all.full.df$SNP, meqtl.all.full.df$CpG_ID, sep = "-")
```

```{r, include=F, eval=T}
eqtm.in.pre  <- "~/bio/code/mpip/dex-stim-human-array/data/integrative/matrixEQTL/"

snp.df.fn <- paste0(eqtm.in.pre, "snp_mtrx.csv")
snp.df    <- fread(snp.df.fn)

methyl.beta.df.fn   <- paste0(eqtm.in.pre, "methyl_beta_mtrx_dex.csv")
methyl.beta.dex.df  <- fread(methyl.beta.df.fn) 

methyl.beta.df.fn   <- paste0(eqtm.in.pre, "methyl_beta_mtrx_veh.csv")
methyl.beta.veh.df  <- fread(methyl.beta.df.fn) 

all(colnames(snp.df)[-1] == colnames(methyl.beta.dex.df)[-1])
all(colnames(snp.df)[-1] == colnames(methyl.beta.veh.df)[-1])
```

<!-- ## Scatter plot and Boxplots for significant meQTLs -->

```{r func-getbetavalues, include = F, eval = T}
GetBetValuesDF <- function(methyl.beta.df, snp.df, selected.qtl, treatment){
  beta.values.df    <- methyl.beta.df[CpG_ID %in% selected.qtl$CpG_ID, -1]
  beta.values.df    <- rbind(beta.values.df, snp.df[SNP %in% selected.qtl$SNP, -1])
  beta.values.df    <- data.frame(t(beta.values.df)) %>% setDT()
  colnames(beta.values.df) <- c("CpG", "SNP")
  beta.values.df$SNP <- as.factor(beta.values.df$SNP)
  beta.values.df[SNP == 0, SNP := "AA"]; beta.values.df[SNP == 1, SNP := "AB"]; beta.values.df[SNP == 2, SNP := "BB"] 
  
  beta.values.df$treatment <- treatment
  beta.values.df
}
```

```{r func-plot-scatterplot, include = F, eval = T}
GetScatterPlot <- function(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05, plot.title = NULL){
  
  cbPalette <- c( "#0072B2", "#009E73", "#E69F00", "#F0E442", "#D55E00", "#CC79A7", "#56B4E9", "#999999")
  
  ggplot(meqtl.all.full.df, aes(x = beta, y = -log10(fdr), shape = treatment, color = treatment)) +
    geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
    geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
    geom_point(alpha = 1.5, size = 1.2) +
    geom_label_repel(data = selected.meqtl,
                     aes(x = beta,
                         y = -log10(fdr),
                         label = meQTL_ID),
                     fontface = 'bold',
                     box.padding = unit(0.35, "lines"),
                     point.padding = unit(0.5, "lines"),
                     segment.color = 'grey50',
                     nudge_x = 0.05, nudge_y = 10, 
                     size = 3) +
    scale_x_continuous(labels = scientific) +
    # scale_y_continuous(trans = trans_reverser('log10')) +
    # labs(title = " ", y = "", x = " ") + 
    theme(legend.position = "bottom", 
          legend.title = element_blank(),
          #  panel.grid.major = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(size = 10),
          axis.title = element_text(size = 8)) +
    labs( x = "Effect size (FC), MatrixEQTL", 
          title = plot.title) +
    scale_colour_manual(values = cbPalette)
}
```

```{r func-plot-boxplot, include = F, eval = T}
GetBoxPlot <- function(beta.values.df, selected.meqtl, fdr.thr = 0.05, plot.labels = c("AA", "AB", "BB"), plot.title = NULL){
  
  cbPalette <- c(  "#009E73", "#E69F00", "#F0E442", "#D55E00", "#CC79A7", "#56B4E9", "#999999")
  
  beta.values.df %>%
    ggplot(aes(y = CpG, x = SNP, fill = treatment)) +
      geom_boxplot(width = 0.2, color = "black") +
      # scale_fill_viridis(discrete = TRUE, alpha = 0.5) +
      scale_x_discrete(labels = plot.labels) +
      theme( panel.background = element_blank(),
             plot.title = element_text(size = 10),
             axis.title = element_text(size = 10),
             axis.text.x = element_text(angle = 0, hjust = 0.5), 
             legend.position = "bottom", 
             legend.title = element_blank()) +
      labs(y = paste0(selected.meqtl$CpG_ID, "\nDNAm beta value"), 
           x = selected.meqtl$SNP,
           title = plot.title) +
    scale_fill_manual(values = cbPalette)
}
```

```{r func-process-and-get-boxplot, include = F, eval = T}
ProcessGetBoxPlot <- function(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, 
                              fdr.thr = 0.05, plot.title = NULL){
  
  beta.values.dex.df <- GetBetValuesDF(methyl.beta.dex.df, snp.df, selected.meqtl, "dex")
  beta.values.veh.df <- GetBetValuesDF(methyl.beta.veh.df, snp.df, selected.meqtl, "veh")
  beta.values.df     <- rbind(beta.values.dex.df, beta.values.veh.df)

  snp.ind.cnt.df <- beta.values.dex.df %>% count(SNP) # beta.values.df[ , .(count = count(SNP)), by = .(SNP, treatment)]
  snp.ind.cnt.df[, label := paste0(SNP, " (n = ", n, ")")]
  
  GetBoxPlot(beta.values.df, selected.meqtl, snp.ind.cnt.df$label, fdr.thr = 0.05, plot.title)
                                               }
```

## DELTA: minimum FDR
***

```{r include = T, eval = T}
# paste0("Scatter plot for the independent significant at FDR <= ", fdr.thr, " meQTLs")
meqtl.id       <- meqtl.all.full.df[treatment == "delta"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        dplyr::select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant across DELTA meQTLs: ",
                     "FC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA: maximum positive FC
***

```{r include = T, eval = T}
meqtl.id <-  meqtl.all.full.df[treatment == "delta"][beta == max(beta), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]
kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "meQTL with the maximum FC across DELTA meQTLs: ",
                     "FC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA: minimum FC
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[treatment == "delta"][beta == min(beta), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "meQTL with the minimum FC across DELTA meQTLs: ",
                     "FC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA: minimum absolute FC
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[treatment == "delta"][abs(beta) == min(abs(beta)), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "meQTL with the minimum absolute FC across DELTA meQTLs: ",
                     "FC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA \\\ (VEH + DEX): min FDR
***

```{r include = T, eval = T}
# summary(meqtl.all.full.df[treatment == "dex"])
only.delta.meqtls <- setdiff(meqtl.all.full.df[treatment == "delta", meQTL_ID], 
                             union(meqtl.all.full.df[treatment == "dex", meQTL_ID],
                                   meqtl.all.full.df[treatment == "veh", meQTL_ID]))
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% only.delta.meqtls][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0("The most significant meQTL across the DELTA meQTLs which don't intersect with eiher VEH or DEX. ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DEX \\\ (VEH + DELTA): min FDR
***

```{r include = T, eval = T}
# summary(meqtl.all.full.df[treatment == "dex"])
only.dex.meqtls <- setdiff(meqtl.all.full.df[treatment == "dex", meQTL_ID], 
                             union(meqtl.all.full.df[treatment == "delta", meQTL_ID],
                                   meqtl.all.full.df[treatment == "veh", meQTL_ID]))
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% only.dex.meqtls][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0("The most significant meQTL across the DEX meQTLs which don't intersect with either VEH or DELTA. ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## VEH \\\ (DEX + DELTA): min FDR
***

```{r include = T, eval = T}
# summary(meqtl.all.full.df[treatment == "dex"])
only.veh.meqtls <- setdiff(meqtl.all.full.df[treatment == "veh", meQTL_ID], 
                           union(meqtl.all.full.df[treatment == "delta", meQTL_ID],
                                 meqtl.all.full.df[treatment == "dex", meQTL_ID]))
selected.meqtl  <- meqtl.all.full.df[meQTL_ID %in% only.veh.meqtls][fdr == min(fdr)] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the VEH meQTLs which don't intersect with either DELTA or DEX. ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## (DEX + DELTA) \\\ VEH: min FDR
***

```{r include = F, eval = T}
# summary(meqtl.all.full.df[treatment == "dex"])
not.veh.meqtls <- setdiff(intersect(meqtl.all.full.df[treatment == "delta", meQTL_ID],
                                    meqtl.all.full.df[treatment == "dex", meQTL_ID]),
                          meqtl.all.full.df[treatment == "veh", meQTL_ID])
```

### min DEX FDR
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% not.veh.meqtls][treatment == "dex"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the DEX meQTLs which are present in DELTA but not in VEH ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

### min DELTA FDR
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% not.veh.meqtls][treatment == "delta"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the DELTA meQTLs which are present in DELTA but not in VEH ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## (DEX + VEH) \\\ DELTA: min FDR
***

```{r include = T, eval = T}
# summary(meqtl.all.full.df[treatment == "dex"])
not.delta.meqtls <- setdiff(intersect(meqtl.all.full.df[treatment == "veh", meQTL_ID],
                                    meqtl.all.full.df[treatment == "dex", meQTL_ID]),
                          meqtl.all.full.df[treatment == "delta", meQTL_ID])
```

### min DEX FDR
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% not.delta.meqtls][treatment == "dex"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the DEX meQTLs which are present in VEH but not in DELTA ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

### min VEH FDR
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% not.delta.meqtls][treatment == "veh"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the VEH meQTLs which are present in DEX but not in DELTA ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## (VEH + DELTA) \\\ DEX: min FDR
***

```{r include = T, eval = T}
not.delta.meqtls <- setdiff(intersect(meqtl.all.full.df[treatment == "veh", meQTL_ID],
                                    meqtl.all.full.df[treatment == "delta", meQTL_ID]),
                          meqtl.all.full.df[treatment == "dex", meQTL_ID])
```

### min VEH FDR
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% not.delta.meqtls][treatment == "veh"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the VEH meQTLs which are present in DELTA but not in DEX ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

### min DELTA FDR
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[meQTL_ID %in% not.delta.meqtls][treatment == "delta"][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id] # beta == max(beta)

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```
```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL across the DELTA meQTLs which are present in VEH but not in DEX ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## Maximum absolute FC 
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[abs(beta) == max(abs(beta)), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "meQTL with the maximum absolute FC across all meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## Maximum FC 
***

```{r include = T, eval = T}
meqtl.id       <- meqtl.all.full.df[beta == max(beta), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID == meqtl.id]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "meQTL with the maximum FC across all meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## meQTL with FC <= 0.02
***

```{r include = T, eval = T}

fc.thr <- 0.02

meqtl.id       <- meqtl.all.full.df[abs(beta) <= fc.thr][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID %in%  meqtl.id ]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL with FC <= ", fc.thr, " across all meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## VEH meQTL with FC <= 0.02
***

```{r include = T, eval = T}

fc.thr <- 0.02

meqtl.id       <- meqtl.all.full.df[treatment == "veh"][abs(beta) <= fc.thr][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID %in%  meqtl.id ]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL with FC <= ", fc.thr, " across VEH meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA meQTL with FC <= 0.02
***

```{r include = T, eval = T}

fc.thr <- 0.02

meqtl.id       <- meqtl.all.full.df[treatment == "delta"][abs(beta) <= fc.thr][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID %in%  meqtl.id ]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL with FC <= ", fc.thr, " across DELTA meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA meQTL with FC <= 0.015
***

```{r include = T, eval = T}

fc.thr <- 0.015

meqtl.id       <- meqtl.all.full.df[treatment == "delta"][abs(beta) <= fc.thr][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID %in%  meqtl.id ]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL with FC <= ", fc.thr, " across DELTA meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```

***
## DELTA meQTL with FC <= 0.01
***

```{r include = T, eval = T}

fc.thr <- 0.01

meqtl.id       <- meqtl.all.full.df[treatment == "delta"][abs(beta) <= fc.thr][fdr == min(fdr), meQTL_ID]
selected.meqtl <- meqtl.all.full.df[meQTL_ID %in%  meqtl.id ]

kable(selected.meqtl %>% 
        select(SNP, CpG_ID, FC = beta, `t-stat`, `p-value`, FDR = fdr, Treatment = treatment) %>% 
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r, eval = T, include = F}
plot.title <- paste0(# "Independent meQTL ", selected.meqtl$meQTL_ID, 
                     "The most significant meQTL with FC <= ", fc.thr, " across DELTA meQTLs: ",
                     "\nFC = ", signif(selected.meqtl$beta, 2), " and FDR = ", signif(selected.meqtl$fdr, 2))
```
```{r out.width = "95%", include = T, eval = T}
GetScatterPlot(meqtl.all.full.df, selected.meqtl, fdr.thr = 0.05,
               plot.title = plot.title)
```
```{r out.width = "95%", include = T, eval = T}
ProcessGetBoxPlot(methyl.beta.veh.df, methyl.beta.dex.df, snp.df, selected.meqtl, fdr.thr = 0.05, 
                  plot.title = plot.title)
```
