***
# __Identification of independent baseline meQTLs: region-wise vs. global__
***

## Methodology

### Global approach

### Region-wise

The number of independent SNPs per cis-region was identified by LD pruning and clumping the SNPs using the “clump” command in PLINK (using distance < __200 kb__ and __r2 > 0.2__ as setting). Each SNP forms a SNP bin, by aggregating all other SNPs into bins by the tag SNP at r2 > 0.2 and distance < 200 kb, such that all SNPs within a given bin were correlated to the current SNP, but to any other SNP. 


```{r, include = F, eval = T}
out.dir.pre  <- "~/bio/code/mpip/dex-stim-human-array/output/data/integrative/matrixEQTL/"

cbPalette <- c("#CC79A7", "#009E73", "#56B4E9", "#0072B2",  "#E69F00", "#F0E442", "#D55E00", "#999999")

meqtl.ind.global.df <- fread(paste0(out.dir.pre, "me-qtl_cis_indp_result_veh_fdr_005.csv"))
meqtl.ind.rw.df     <- fread(paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_indp_rw_veh_fdr_005.csv"))
```

```{r include = F, eval = F}
summary(meqtl.ind.global.df)
```

```{r include = F, eval = F}
summary(meqtl.ind.rw.df)
```

```{r include = F, eval = T}
meqtl.ind.rw.df[["meQTL_ID"]] <- paste(meqtl.ind.rw.df$SNP, meqtl.ind.rw.df$CpG_ID, sep = "-")

meqtl.ind.global.df[["meQTL_ID"]] <- paste(meqtl.ind.global.df$SNP, meqtl.ind.global.df$CpG_ID, sep = "-")

meqtl.full.df <- rbind(meqtl.ind.global.df[, approach := "global"],
                             meqtl.ind.rw.df[, approach := "region-wise"])

meqtl.full.df <- setDT(meqtl.full.df)
```

## Upset plots{.tabset .tabset-fade .tabset-pills}

### meQTL level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.meqtls <- list(global = meqtl.full.df[approach == "global", meQTL_ID], 
                      region_wise = meqtl.full.df[approach == "region-wise", meQTL_ID])

upset(fromList(meqtls.meqtls), 
      sets = c("global", "region_wise"),
      nsets = 2, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, meQTLs", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = cbPalette[1:2], 
      order.by = "freq")
```

### SNP level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.snps <-  list(global = meqtl.full.df[approach == "global", SNP], 
                     region_wise = meqtl.full.df[approach == "region-wise", SNP])

upset(fromList(meqtls.snps), 
      sets = c("global", "region_wise"),
      nsets = 2, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, SNPs", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = cbPalette[1:2], 
      order.by = "freq")
```

### CpG level

```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.cpgs <-  list(global = meqtl.full.df[approach == "global", CpG_ID], 
                     region_wise = meqtl.full.df[approach == "region-wise", CpG_ID])

upset(fromList(meqtls.cpgs), 
      sets = c("global", "region_wise"),
      nsets = 2, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, CpGs", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = cbPalette[1:2], 
      order.by = "freq")
```
## Scatter plot

```{r include = F, eval = T}
source("~/bio/code/mpip/dex-stim-human-array/code/integrative/eqtl/04_me-qtl_analysis/util.R")
```

```{r out.width = "95%", include = F, eval = F}
fdr.thr <- 0.05
plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")
GetScatterPlot2(meqtl.all.full.df, plot.title = plot.title)
```

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05

plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")

ggplot(meqtl.full.df[order(`p-value`, approach),], 
       aes(x = beta, y = -log10(fdr), 
           shape = approach, color = approach)) +
  geom_point(alpha = 1.5, size = 2) +
  scale_x_continuous(labels = scientific) +
  geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
  geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "Effect size (FC), MatrixEQTL",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```


## Significance comparison

```{r include = F, eval = T}
meqtl.delta.only.global.df <- meqtl.ind.global.df[!(meQTL_ID %in% meqtl.ind.rw.df$meQTL_ID), ]

meqtl.delta.only.rw.df <- meqtl.ind.rw.df[!(meQTL_ID %in% meqtl.ind.global.df$meQTL_ID), ]

meqtl.unique.df <- rbind( meqtl.delta.only.rw.df, meqtl.delta.only.global.df)

meqtl.union.df <- meqtl.ind.global.df[(meQTL_ID %in% meqtl.ind.rw.df$meQTL_ID), ] %>% mutate(approach = "global + region-wise")

```

```{r out.width = "95%", include = T, eval = T}
plot.title <- paste0("Distribution of p-values of unique and intersected baseline meQTLs obtained by global and region-wise approaches.")

ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
       aes(x = approach, y = -log10(fdr), 
           shape = approach, color = approach)) +
  geom_boxplot(position=position_dodge(1)) +
  scale_y_continuous(labels = scientific) +
  theme(legend.position = "none", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```


```{r out.width = "95%", include = T, eval = T}
plot.title <- paste0("Distribution of beta-values of unique and intersected baseline meQTLs obtained by global and region-wise approaches.")

ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
       aes(x = approach, y = beta, 
           shape = approach, color = approach)) +
  geom_boxplot(position=position_dodge(1)) +
  scale_y_continuous(labels = scientific) +
  geom_hline(yintercept = 0, colour = "#990000", linetype = "dashed") +
  theme(legend.position = "none", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05

plot.title <- paste0("Volcano plot for unique and intersected baseline meQTLs obtained by global and region-wise approaches.")

ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
       aes(x = beta, y = -log10(fdr), 
           shape = approach, color = approach)) +
  geom_point(alpha = 1.5, size = 2) +
  scale_x_continuous(labels = scientific) +
  geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
  geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "Effect size (FC), MatrixEQTL",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```

```{r include = T, eval = T}
meqtl.unique.df <- meqtl.unique.df[order(meqtl.unique.df$fdr), ]

selected.meqtl <- head(meqtl.unique.df, 10)

kable(selected.meqtl %>%
        select(meQTL = meQTL_ID, SNP, CpG = CpG_ID, FC = beta, FDR = fdr, `p-value`, `t-stat`, Approach = approach) %>%
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r  out.width = "95%", include = T, eval = T}
 ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
        aes(x = beta, y = -log10(fdr), shape = approach, color = approach)) +
    geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
    geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
    geom_point(alpha = 1.5, size = 1.2) +
    geom_label_repel(data = selected.meqtl,
                     aes(x = beta,
                         y = -log10(fdr),
                         label = meQTL_ID),
                     fontface = 'bold',
                     box.padding = unit(0.35, "lines"),
                     point.padding = unit(0.5, "lines"),
                     segment.color = 'grey50',
                     nudge_x = -0.009, 
                     nudge_y = -1.5, 
                     size = 2) +
    scale_x_continuous(labels = scientific) +
    # scale_y_continuous(trans = trans_reverser('log10')) +
    # labs(title = " ", y = "", x = " ") + 
    theme(legend.position = "bottom", 
          legend.title = element_blank(),
          #  panel.grid.major = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(size = 10),
          axis.title = element_text(size = 8)) +
    labs( x = "Effect size (FC), MatrixEQTL", 
          title = plot.title) +
    scale_colour_manual(values = cbPalette)
```
