---
title: "cis-meQTLs Analysis"
subtitle: "Dexamethasone-Stimulated Human Array Project"
author: 
  - name: Anastasiia Hryhorzhevska
    email: anastasiia_hry@psych.mpg.de
    url: https://github.com/ahryho
    affiliation: Max Planck Institute of Psychiatry
    affiliation_url: https://www.psych.mpg.de/2664393/medizinische-genomforschung
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    css: style.css
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
options(digits = 4, width = 100, stringsAsFactors = T)
knitr::opts_chunk$set(echo = TRUE,
                      tidy.opts = list(width.cutoff = 100),
                      tidy=TRUE,
                      fig.pos = "H",
                      dpi = 200,
                      warning = FALSE, 
                      message = FALSE, 
                      cache = TRUE, 
                      cache.lazy = FALSE)

# define DT as the standard printing method for data.frame
library(knitr)
library(DT)
knit_print.data.frame <- function(x, ...) {
  knit_print(DT::datatable(x,
                           filter="top",
                           rownames = FALSE,
                           extensions = "FixedColumns",
                           options = list(
                             scrollX = TRUE,
                             pageLength = 5
                             
                           )), ...)
}

registerS3method("knit_print", "data.frame", knit_print.data.frame)

library(data.table)
library(dplyr)

library(ggplot2)
library(ggrepel)
library(scales)
cbPalette <- c("#CC79A7", "#009E73", "#56B4E9", "#0072B2",  "#E69F00", "#F0E442", "#D55E00", "#999999")

```

***
# __Identification of independent delta meQTLs: region-wise vs. global__
***

## Methodology

### Global approach

### Region-wise

The number of independent SNPs per cis-region was identified by LD pruning and clumping the SNPs using the “clump” command in PLINK (using distance < __200 kb__ and __r2 > 0.2__ as setting). Each SNP forms a SNP bin, by aggregating all other SNPs into bins by the tag SNP at r2 > 0.2 and distance < 200 kb, such that all SNPs within a given bin were correlated to the current SNP, but to any other SNP. 


```{r, include = F, eval = T}
out.dir.pre  <- "~/bio/code/mpip/dex-stim-human-array/output/data/integrative/matrixEQTL/"

meqtl.delta.ind.global.df <- fread(paste0(out.dir.pre, "me-qtl_cis_indp_result_delta_fdr_005.csv"))
meqtl.delta.ind.rw.df     <- fread(paste0(out.dir.pre, "region_wise_independent_snps/me-qtl_cis_indp_rw_delta_fdr_005.csv"))
```

```{r include = F, eval = F}
summary(meqtl.delta.ind.global.df)
```

```{r include = F, eval = F}
summary(meqtl.delta.ind.rw.df)
```

```{r include = F, eval = T}
meqtl.delta.ind.rw.df[["meQTL_ID"]] <- paste(meqtl.delta.ind.rw.df$SNP, meqtl.delta.ind.rw.df$CpG_ID, sep = "-")

meqtl.delta.ind.global.df[["meQTL_ID"]] <- paste(meqtl.delta.ind.global.df$SNP, meqtl.delta.ind.global.df$CpG_ID, sep = "-")

meqtl.delta.full.df <- rbind(meqtl.delta.ind.global.df[, approach := "global"],
                             meqtl.delta.ind.rw.df[, approach := "region-wise"])

meqtl.delta.full.df <- setDT(meqtl.delta.full.df)
```

## Upset plots{.tabset .tabset-fade .tabset-pills}

### meQTL level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.meqtls <- list(global = meqtl.delta.full.df[approach == "global", meQTL_ID], 
                      region_wise = meqtl.delta.full.df[approach == "region-wise", meQTL_ID])

upset(fromList(meqtls.meqtls), 
      sets = c("global", "region_wise"),
      nsets = 2, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, meQTLs", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = cbPalette[1:2], 
      order.by = "freq")
```

### SNP level
  
```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.snps <-  list(global = meqtl.delta.full.df[approach == "global", SNP], 
                     region_wise = meqtl.delta.full.df[approach == "region-wise", SNP])

upset(fromList(meqtls.snps), 
      sets = c("global", "region_wise"),
      nsets = 2, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, SNPs", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = cbPalette[1:2], 
      order.by = "freq")
```

### CpG level

```{r out.width = "95%", include = T, eval = T}
library(UpSetR)

meqtls.cpgs <-  list(global = meqtl.delta.full.df[approach == "global", CpG_ID], 
                     region_wise = meqtl.delta.full.df[approach == "region-wise", CpG_ID])

upset(fromList(meqtls.cpgs), 
      sets = c("global", "region_wise"),
      nsets = 2, 
      nintersects = 20, 
      mainbar.y.label = "Number of intersections, CpGs", 
      text.scale = 1, 
      keep.order = T,
      sets.bar.color = cbPalette[1:2], 
      order.by = "freq")
```
## Scatter plot

```{r include = F, eval = T}
source("~/bio/code/mpip/dex-stim-human-array/code/integrative/eqtl/04_me-qtl_analysis/util.R")
```

```{r out.width = "95%", include = F, eval = F}
fdr.thr <- 0.05
plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")
GetScatterPlot2(meqtl.all.full.df, plot.title = plot.title)
```

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05

plot.title <- paste0("Independent and significant at FDR <= ", fdr.thr, " meQTLs.",
                     "\nResult of MatrixEQTL")

ggplot(meqtl.delta.full.df[order(`p-value`, approach),], 
       aes(x = beta, y = -log10(fdr), 
           shape = approach, color = approach)) +
  geom_point(alpha = 1.5, size = 2) +
  scale_x_continuous(labels = scientific) +
  geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
  geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "Effect size (FC), MatrixEQTL",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```


## Significance comparison

```{r include = F, eval = T}
meqtl.delta.only.global.df <- meqtl.delta.ind.global.df[!(meQTL_ID %in% meqtl.delta.ind.rw.df$meQTL_ID), ]

meqtl.delta.only.rw.df <- meqtl.delta.ind.rw.df[!(meQTL_ID %in% meqtl.delta.ind.global.df$meQTL_ID), ]

meqtl.unique.df <- rbind( meqtl.delta.only.rw.df, meqtl.delta.only.global.df)

meqtl.union.df <- meqtl.delta.ind.global.df[(meQTL_ID %in% meqtl.delta.ind.rw.df$meQTL_ID), ] %>% mutate(approach = "global + region-wise")

```

```{r out.width = "95%", include = T, eval = T}
plot.title <- paste0("Distribution of p-values of unique and intersected delta meQTLs obtained by global and region-wise approaches.")

ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
       aes(x = approach, y = -log10(fdr), 
           shape = approach, color = approach)) +
  geom_boxplot(position=position_dodge(1)) +
  scale_y_continuous(labels = scientific) +
  theme(legend.position = "none", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```


```{r out.width = "95%", include = T, eval = T}
plot.title <- paste0("Distribution of beta-values of unique and intersected delta meQTLs obtained by global and region-wise approaches.")

ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
       aes(x = approach, y = beta, 
           shape = approach, color = approach)) +
  geom_boxplot(position=position_dodge(1)) +
  scale_y_continuous(labels = scientific) +
  geom_hline(yintercept = 0, colour = "#990000", linetype = "dashed") +
  theme(legend.position = "none", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```

```{r out.width = "95%", include = T, eval = T}
fdr.thr <- 0.05

plot.title <- paste0("Volcano plot for unique and intersected delta meQTLs obtained by global and region-wise approaches.")

ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
       aes(x = beta, y = -log10(fdr), 
           shape = approach, color = approach)) +
  geom_point(alpha = 1.5, size = 2) +
  scale_x_continuous(labels = scientific) +
  geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
  geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  labs(x = "Effect size (FC), MatrixEQTL",
       title = plot.title) +
  scale_colour_manual(values = cbPalette)
```

```{r include = T, eval = T}
meqtl.unique.df <- meqtl.unique.df[order(meqtl.unique.df$fdr), ]

selected.meqtl <- head(meqtl.unique.df, 10)

kable(selected.meqtl %>%
        select(meQTL = meQTL_ID, SNP, CpG = CpG_ID, FC = beta, FDR = fdr, `p-value`, `t-stat`, Approach = approach) %>%
        mutate_if(is.numeric, funs(as.character(signif(., 3)))))
```

```{r  out.width = "95%", include = T, eval = T}
 ggplot(rbind(meqtl.union.df, meqtl.unique.df), 
        aes(x = beta, y = -log10(fdr), shape = approach, color = approach)) +
    geom_vline(xintercept = 0, colour = "#990000", linetype = "dashed") + 
    geom_hline(yintercept = -log10(fdr.thr), colour = "#990000", linetype = "dashed") +
    geom_point(alpha = 1.5, size = 1.2) +
    geom_label_repel(data = selected.meqtl,
                     aes(x = beta,
                         y = -log10(fdr),
                         label = meQTL_ID),
                     fontface = 'bold',
                     box.padding = unit(0.35, "lines"),
                     point.padding = unit(0.5, "lines"),
                     segment.color = 'grey50',
                     nudge_x = -0.009, 
                     nudge_y = -1.5, 
                     size = 2) +
    scale_x_continuous(labels = scientific) +
    # scale_y_continuous(trans = trans_reverser('log10')) +
    # labs(title = " ", y = "", x = " ") + 
    theme(legend.position = "bottom", 
          legend.title = element_blank(),
          #  panel.grid.major = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(size = 10),
          axis.title = element_text(size = 8)) +
    labs( x = "Effect size (FC), MatrixEQTL", 
          title = plot.title) +
    scale_colour_manual(values = cbPalette)
```
